### 2021-04-22 paper_analysis.Rmd
update

---
title: "Paper analysis"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Descriptivos

## Univariado
```{r echo=FALSE, results='asis',message=F}
pacman::p_load(dplyr, sjlabelled, summarytools, ggplot2, lme4, texreg, stargazer, car)
load("input/data-proc/data.rda")
st_css()
basereg <- data %>% 
  dplyr::select(
    igualdad.est,
    esp_publico.est,
    esp_privado.est,
    igualdad.pad,
    esp_publico.pad,
    esp_privado.pad,
    ap_discusion,
    c_civic,
    Sexo,
    num_fem,
    Educacion,
    quintiles_ingresos_pc_factor,
    Libros_apod_rec,
    dependencia,
    RBD,
    REGION
  ) %>% 
  ungroup() %>% 
  na.omit()

print(dfSummary(basereg,
               plain.ascii = FALSE,
               style = "grid",
               tmp.img.dir = "/tmp",
               graph.magnif = 0.75,
               headings = F,  # encabezado
               varnumbers = F, # num variable
               labels.col = T, # etiquetas
               na.col = F,    # missing
               graph.col = T, # plot
               valid.col = T, # n valido  
               # col.widths = c(5,5,10,10,10,10)
               ),
     method = "render")
```

## Bivariado
```{r}
pacman::p_load(summarytools,haven,sjPlot,dplyr)
load("input/data-proc/data.rda")
basereg <- data %>% 
  dplyr::select(
    igualdad.est,
    esp_publico.est,
    esp_privado.est,
    Sexo,
    ap_discusion,
    mean_discusion,
    num_fem,
    RBD,
    igualdad.pad,
    esp_publico.pad,
    esp_privado.pad,
    Educacion,
    Libros_apod,
    Libros_apod_rec,
    quintiles_ingresos_pc_factor,
    dependencia,
    REGION,
    c_civic,
  ) %>% na.omit()
```

# Modelos